---
title: "Assignment8"
author: "NguyenManhHung_22BI13183"
date: "2025-04-01"
output: html_document
---
```{r}
library(clusterProfiler)
library(org.Hs.eg.db)
library(enrichplot)
library(fgsea)
library(DOSE)
library(ggplot2)
library(stats)
```
## QUESTION 1: What does each dot in the dot plots above represent?

Answer: - Each dot represents a Gene Ontology term with the dot size is the number of genes enriched in that term.The dot color is the p-adjected value, with red indicates more significant enrichment and blue indicates less significant one.The x-axis is the proportion of genes from the input list associated with a particular GO term.

## QUESTION 2: In the same gene list above, create a dot plot showing the top 10 most enriched GO Cellular Components term by Over-representation Analysis. Which GO Cellular Components term is the most enriched?
```{r}
# Load the dataset
deg <- readRDS("Breast-cancer_DEG.RDS")

# using row as ids
deg$entrez <- as.character(rownames(deg))

# rename the column
colnames(deg)[colnames(deg) == "Log2FC"] <- "log2FoldChange"

# Filter for |log2FC| > 1
deg_filtered <- deg[abs(deg$log2FoldChange) > 1, ]

#find valid ids
valid_entrez <- keys(org.Hs.eg.db, keytype = "ENTREZID")

# Keep only valid IDs
deg_filtered <- deg_filtered[deg_filtered$entrez %in% valid_entrez, ]
entrez_ids <- deg_filtered$entrez
```

```{r}
# ORA for Cellular Component
ora_cc <- enrichGO(
  gene          = entrez_ids,
  OrgDb         = org.Hs.eg.db,
  keyType       = "ENTREZID",
  ont           = "CC",
  pAdjustMethod = "BH",
  qvalueCutoff  = 0.05,
  pvalueCutoff  = 0.05,
  readable      = TRUE
)

# Plot
dotplot(ora_cc, showCategory = 10, title = "Top 10 GO Cellular Component Terms (ORA)")
```
```{r}
# Most enriched Cellular Component term (ORA)
top <- ora_cc@result[order(ora_cc@result$p.adjust), ][1, ]
name <- top$Description
p_adj <- top$p.adjust
cat('Most enriched CC term:', name , '\n')
cat('p-value: ',p_adj,"\n")
```
## QUESTION 3: Similarly, using the same gene list, create a dot plot showing the top 10 most enriched Biological Processes term by Gene Set Enrichment Analysis. Which BP Cellular Components term is the most enriched?


```{r}
# ORA for Biological Process
ora_bp <- enrichGO(
  gene          = entrez_ids,
  OrgDb         = org.Hs.eg.db,
  keyType       = "ENTREZID",
  ont           = "BP",
  pAdjustMethod = "BH",
  qvalueCutoff  = 0.05,
  pvalueCutoff  = 0.05,
  readable      = TRUE
)

# Plot
dotplot(ora_bp, showCategory = 10, title = "Top 10 GO Biological Process Terms (ORA)")


```
```{r}
#get the most riched
top <- ora_bp@result[order(ora_bp@result$p.adjust), ][1, ]
  
# Extract name and adjusted p-value
name <- top$Description
p_adj <- top$p.adjust

# Print results
cat("Most enriched BP term:", name, "\n")
cat("p-value:", p_adj, "\n")

```



